Rust for Scientific Computing

### 2025-08-25 main.tex
fixup! cleanup
add more comments and notes for included crates

ploc\\
A parallelized algorithm library for localization and clustering tasks. Focuses on scalable
implementations of clustering approaches. Intended for large datasets in scientific applications.

@@

@@

ploc\\
Efficient point location with ploc.

NPB-Rust\\
NAS parallel benchmarks in Rust.

@@

@@

Juice\\
Juice your simulations: what science can learn from game development.

eVaiutilities\\
Research data management post eVai.

@@

Stochastic\\
Stochastic calculus in Rust.

Exponential\\
Exponential time integration for stiff systems in Rust.

Rust\\
Rust for chip design algorithms and EDA software.

Using\\
Using Rust to mitigate breaking changes in scientific software.

The\\
The pleasure and pain of developing cross-platform CPU vector code in Rust.

@@

Rust\\
Rust is RAD and this is why.

ploc\\
Efficient point location with ploc.

@@

@@

@@

Rootfinders\\
Rootfinders for Rust.

@@

@@

@@

@@

Cargo\\
Cargo +GPU build: An early outlook.

@@

\paragraph{FAANG Companies}

>>>>

ploc\\
A parallelized algorithm library for localization and clustering tasks. Focuses on scalable
implementations of clustering approaches. Intended for large datasets in scientific applications.
\url{https://dotblocks.com/}

@@

stochastic-rs\\
stochastic-rs is a high-performance Rust library for simulating and analyzing
stochastic processes. Designed for applications in quantitative finance, AI training, and
statistical modeling, it provides efficient tools to generate synthetic data and analyze complex
stochastic systems.

@@

NPB-Rust\\
NAS parallel benchmarks in Rust. \cite{martins2025npbrustnasparallelbenchmarks}

@@

ORMATEX\\
Oak Ridge MATrix EXponential tools - methods to compute the matrix exponential

orbweaver-rs\\
A fast R library for working with Nodes in a graph.

triple-r\\
Recycle, Reuse, Reduce. triple-r is a high-performance Rust library that provides
wrappers around standard library collections to enable the reuse of their memory allocations. By
recycling the underlying memory of collections like HashMap and Vec, triple-r helps reduce
allocation overhead in performance-critical applications.

faucet (maybe remove; does this fit the scope?)\\
Scale, deploy and route Plumber APIs and Shiny
applications with ease and efficiency.

qoqo\\
A toolkit to represent quantum circuits by HQS Quantum Simulations.

@@

eVaiutilities\\
A data management software for the analysis of the eVai output files. It allows the
analysis of the genomic variants further such as analyzing the multiple genomic annotated variants,
reference and alternate allele, enabling coordinate search, coordinate search with specified
variants and annotation search across a large number of population.

@@

qhull-rs\\
Qhull computes the convex hull, Delaunay triangulation, Voronoi diagram, halfspace
intersection about a point, furthest-site Delaunay triangulation, and furthest-site Voronoi
diagram.

HULK (maybe out of scope?)\\
The robot control program and associated tools of the RoboCup SPL team
HULKs. \url{https://hulks.de/hulk/} "Rust’s safety and performance features benefit real-time
robotics, and the challenges of integrating it into an ecosystem traditionally dominated by C++ and
Python."

burn\\
Burn is a next generation Deep Learning Framework that doesn't compromise on flexibility,
efficiency and portability. \url{https://github.com/tracel-ai/burn}\\
Commercial

@@

@@

oxidicom\\
A high-performance DICOM receiver for the ChRIS backend (CUBE).
\url{https://chrisproject.org/}

@@

cfsem\\
Quasi-steady electromagnetics including filamentized approximations, Biot-Savart, and
Grad-Shafranov.\\
\url{https://cfs.energy/} Commercial

@@

Argmin\\
Rootfinders for Rust.

@@

Symbolica\\
Symbolica is a blazing fast computer algebra system for Python and Rust, born of a need
to push the boundaries of computations in science and enterprise.

@@

rsrs (no crate, just github repo)\\
An implementation of the randomised strong recursive
skeletonisation

bevy\\
Bevy is a refreshingly simple data-driven game engine built in Rust.

Polars\\
Polars is a DataFrame library for Rust. It is based on Apache Arrow’s memory model. Apache
Arrow provides very cache efficient columnar data structures and is becoming the defacto standard
for columnar data.

@@

Roxido\\
Developing Rust-based R packages using the Roxido framework.

@@

Biomechanics Foundation\\
Back to basics: rebuilding computational biomechanics crate by crate.
Many crates (maybe just cite one crate):
\begin{itemize}
    \item chiron - Graphical and command-line interface tools for Biomechanics Foundation
    \item gcv\_spline - Generalized Cross-Validated Splines for Interpolation and Derivation in Pure Rust
    \item onager - Featherstone rigid-body physics library built in Rust
    \item c3dio - A c3d parser, writer and editor written in Rust.
    \item bevy\_c3d - A .c3d asset loader plugin for the Bevy engine.
    \item Mannequin - Methods for kinematics and rigging, and dynamics in biomechanics applications and
          animation
    \item hominem - Biomechanical muscle models for simulating human motion
\end{itemize}
\url{https://biomechanicsfoundation.org/}
\url{https://github.com/biomechanics-foundation}

\#[offload] (feature draft)\\
New rustc feature called \#[offload], which allows running almost arbitrary Rust code on
co-processors like a GPU, TPU, or IPU
\url{https://github.com/rust-lang/rust/pull/145768}

Bempp-rs\\
Writing a grid library for finite and boundary element methods. \cite{Witherden2015}

@@
